---
title: ""
output:
  html_document:
    df_print: paged
  html_notebook:
    highlight: zenburn
editor_options:
  chunk_output_type: console
---


```{r  global_options}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE)
```

### COVID-19 in North Korea

```{r  results='hide'}
library(readxl)
library(ggplot2)

dat <- read_xlsx("data/covid19_daily.xlsx")
names(dat) <- c("date", "symptomatic", "recovered", "deaths", "cumul_symptomatic", "cumul_recovered", "cumul_isolated", "cumul_deaths")
dat <- dat[3:(nrow(dat)-1),]
library(dplyr)
dat %>% 
  mutate(across(2:8, as.integer),
         newdate = seq(from=as.Date("2022-05-13"), 
                       to=as.Date("2022-08-10"), by="day"), 
         cfr = cumul_deaths / cumul_symptomatic) -> dat

# d <- dat[,"newdate", ]
# sum((dat$cumul_symptomatic - dat$cumul_recovered - dat$cumul_isolated) != 0)
# tstamp <- format(Sys.time(), "%Y%m%d")
# saveRDS(dat, paste0("outputs/NorthKorea_dat_", tstamp, ".rds"))
```

- 신규 유열자 (new symptomatic cases) were first repoorted on `r format(dat$newdate[1], "%B %d, %Y")` ($n$ = `r dat$symptomatic[1]`).
- There were, however, `r dat$cumul_symptomatic[1]` symptomatic cases already on `r format(dat$newdate[1], "%B %d, %Y")`.
- In total, `r dat$cumul_symptomatic[nrow(dat)]` cases were reported `r format(dat$newdate[1], "%B %d, %Y")` through `r format(dat$newdate[nrow(dat)], "%B %d, %Y")`
- Case fatality ratio was very low: CFR = `r format(round(dat$cfr[nrow(dat)], digits=8), scientific=FALSE)`

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(tidyr)

dat %>%  select(newdate, symptomatic, cumul_symptomatic) %>%
  pivot_longer(cols=c(2,3), names_to = "daily_cumul") -> d

plt <- ggplot(d, aes(newdate, value))+
  geom_line()+
  geom_point()+
  labs(x="", y="")+
  ggtitle("Symptomatic cases")+
  facet_wrap(~daily_cumul, nrow=2, scales="free_y")

plt
# Alene, Muluneh, Leltework Yismaw, Moges Agazhe Assemie, Daniel Bekele Ketema, Wodaje Gietaneh, and Tilahun Yemanu Birhan. “Serial Interval and Incubation Period of COVID-19: A Systematic Review and Meta-Analysis.” BMC Infectious Diseases 21, no. 1 (December 2021): 1–9. https://doi.org/10.1186/s12879-021-05950-x.

#Abstract: ... mean serial interval of COVID-19 was 5.2 (95%CI: 4.9–5.5) days
SI <- 5.2
SI_std <- (5.5-5.2)/1.96
```

### $R_t$ analysis

- EpiEstim package
- 신규 유열자 was used as daily incidence
- Serial interval was was assumed to be `r SI`.
- $R_t$ decreased over time and almost never reached 1

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(EpiEstim)
res <- estimate_R(dat$symptomatic, 
                  method="parametric_si",
                  config=make_config(list(mean_si=SI, std_si=SI_std)))

plot(res)
```


### Particle filter  - North Korea data
Particle filtering - particle_filter
Expect errors since backward sampling is not in place
North Korea data
```{r}
library(pfilterCOVID)
set.seed(1)
theme_set(theme_bw())
tstamp <- "20220105"
I0 <- 100
# data from the pfilterCOVID package
pdat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
# # determine the data type to model (infection vs. confirmation)
# dtype <- "confirmation"
# d <- dat$ode[, c("time", "daily_infected")]
# if (dtype == "confirmation") {
#   d <- dat$ode[, c("time", "daily_confirmed")]
# }
# 
# d <- dat$ode[, c("time", "daily_infected")]
# if (dtype == "confirmation") {
#   d <- dat$ode[, c("time", "daily_confirmed")]
# }
# https://www.google.com/search?q=%EB%B6%81%ED%95%9C%EC%9D%B8%EA%B5%AC&rlz=1C1GCEU_en-GBKR1039KR1040&oq=%EB%B6%81%ED%95%9C%EC%9D%B8%EA%B5%AC&aqs=chrome..69i57j46i199i465i512j0i512j46i175i199i512j0i512l5.3544j0j7&sourceid=chrome&ie=UTF-8
# Pop size in North Korea = 25.97*1e6
Y0 <- c(S=25970000-2e4, E=0, P=0, A=0, I=2e4, R=0, CE=0, CI=0, CR=0)

# dtype <- "symptom onset"
dtype <- "infection"
d <- data.frame(time = as.numeric(0:(nrow(dat)-1)), daily_infected = as.numeric(dat$symptomatic))
d <- d[1:78,]

pdat$params$time_dep_Rt <- FALSE # this must be set FALSE to produce an estimate
params <- pdat$params
#standard deviation of the prior distribution
params[["betavol"]] <- 0.8 

tbegin <- Sys.time()

pf <- pfilter(
  params = params,
  y = Y0,
  data = d,
  data_type = dtype,
  npart = 1e3,
  tend = nrow(d),
  dt = 0.1,
  error_pdf = "negbin",
  # error_pdf = "pois",
  negbin_size = 5,
  systematic_resampling = FALSE,
  backward_sampling = TRUE,
  stoch = FALSE)

Sys.time() - tbegin

# Rt <- readRDS("outputs/Rt_default.rds")
epsilon <- params[["epsilon"]]
delta <- params[["delta"]]
gamma <- params[["gamma"]]
bp <- params[["bp"]] # relative infectiousness of pre-symptomatic state
ba <- params[["ba"]] # relative infectiousness of asymptomatic state
fa <- params[["fa"]] # fraction of asymptomatic state

durP <- (1 / delta - 1 / epsilon)
durI <- (1 / gamma)
R0_dur <- ((1 - fa) + fa * ba) * durI + bp * durP

med <- apply(pf$beta_filtered, 1, quantile, probs = c(0.5))
lb <- apply(pf$beta_filtered, 1, quantile, probs = c(0.025))
ub <- apply(pf$beta_filtered, 1, quantile, probs = c(0.975))

plot(d$time, med * R0_dur, type="l", lwd=2, col=1, ylim=c(0,100))
lines(d$time, ub * R0_dur, lwd=1, col=3)
lines(d$time, lb * R0_dur, lwd=1, col=3)
abline(h = 1, ltype=2)
# plot(dat$time + 2, dat$daily_Rt, type="l", ylim=c(0, max(upper * R0_dur)))
# plot(dat$ode$time + 2, dat$ode$daily_Rt, type="l")
# lines(med * R0_dur, lwd=2, col=3)
df <- cbind(data.frame(date = seq(dat$newdate[1], length.out = nrow(d), by="day")),
            t(apply(pf$beta_filtered, 1, quantile, probs = c(0.025, 0.5, 0.975))))
df$`2.5%` <- df$`2.5%` * R0_dur
df$`97.5%` <- df$`97.5%` * R0_dur
df$`50%`<- df$`50%` * R0_dur
df$trace <- pf$trace$beta * R0_dur

plt <- ggplot(df, aes(x = date)) +
    geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "steelblue", alpha = 0.5) +
    # geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
    geom_line(aes(y = `50%`), color = "steelblue", size = 1.2) +
  geom_line(aes(y = trace), color = "darkgreen", size = 1.2) +
    geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dashed")+
    labs(title = "", y = expression(italic(R)[italic(t)]), x = "")
plt
ggsave("plots/daily_Rt.png", plot=plt)


daily_inf <- t(apply(pf$latent_var_filtered[,,7], 2, quantile, probs = c(0.025, 0.5, 0.975)))
df <- cbind(data.frame(date = seq(dat$newdate[1], length.out = nrow(d), by="day")), 
            daily_inf, trace = pf$trace$CE, observed = d$daily_infected)

plt <- ggplot(df, aes(x = date)) +
    geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "steelblue", alpha = 0.5) +
    # geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
    geom_line(aes(y = `50%`), color = "steelblue", size = 1.2) +
    geom_line(aes(y = trace), color = "darkgreen", size = 1.2) +
    geom_point(aes(y = observed), color = "darkred", size = 1)+
    labs(title = "", y = "New symptomatic cases", x = "")
plt

ggsave("plots/daily_case_model_obs.png", plot=plt)
```



## Particle filtering with backward sampling
```{r}
library(pfilterCOVID)
set.seed(1)
tstamp <- "20220105"
I0 <- 100
# data from the pfilterCOVID package
pdat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
# https://www.google.com/search?q=%EB%B6%81%ED%95%9C%EC%9D%B8%EA%B5%AC&rlz=1C1GCEU_en-GBKR1039KR1040&oq=%EB%B6%81%ED%95%9C%EC%9D%B8%EA%B5%AC&aqs=chrome..69i57j46i199i465i512j0i512j46i175i199i512j0i512l5.3544j0j7&sourceid=chrome&ie=UTF-8
# pop size in North Korea = 25.97*1e6

pdat$params$Y0 <- c(S=25970000-2e4, E=0, P=0, A=0, I=2e4, R=0, CE=0, CI=0, CR=0)
usethis::use_data(pdat$params, overwrite = TRUE) #### if asked to overwrite

# dtype <- "symptom onset"
dtype <- "infection"
d <- data.frame(time = as.numeric(0:(nrow(dat)-1)), daily_infected = as.numeric(dat$symptomatic))
d <- d[1:78,]

pdat$params$time_dep_Rt <- FALSE # this must be set FALSE to produce an estimate
params <- pdat$params
#standard deviation of the prior distribution
params[["betavol"]] <- 0.8 

tbegin <- Sys.time()


# increase the number to improve accuracy
nrep <- 1e2
npart <- 1e3
dt <- 0.2
library(parallel)
library(doParallel)

ncores <- detectCores() 
dtype <- "confirmation"
# ODE with perfect observation
d <- data.frame(date = dat$time,
               daily_infected = round(dat$ode$daily_infected))

if (dtype == "confirmation") {
  d <- data.frame(date = dat$time,
                  daily_confirmed = round(dat$ode$daily_confirmed))
}

set.seed(42)
cl <- makeCluster(getOption("cl.cores", 2))
doParallel::registerDoParallel(cl)

pf <- foreach(i = 1:nrep, .packages = c("pfilterCOVID"), .inorder = F) %dopar% {
 extract_trace(params = params,
                    y = params$Y0,
                    data = d,
                    data_type = dtype,
                    npart = npart,
                    tend = nrow(d),
                    dt = dt,
                    error_pdf = "pois",
                    negbin_size = 15,
                    backward_sampling = TRUE,
                    stoch = FALSE)
}



parallel::stopCluster(cl)
  
parset_chr <- paste0(dtype, "_I0=", params$I0, "_npart=", npart,
           "_nrep=", nrep, "_dt=", dt, "_", tstamp)
# saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
  
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
  
df <- cbind(Rt_quantile, d)
col_fill <- "#1F618D"
col_dat <- "grey70"
df$Rt <- dat$ode$daily_Rt

plt <- ggplot(df, aes(x = date)) +
    geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
    geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
    geom_line(aes(y = `50%`), color = col_fill, size = 1.2) +
    geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
    geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
    labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
plt
```

### Run particle filter
```{r}
# initial values for state variables
# https://www.google.com/search?q=%EB%B6%81%ED%95%9C%EC%9D%B8%EA%B5%AC&rlz=1C1GCEU_en-GBKR1039KR1040&oq=%EB%B6%81%ED%95%9C%EC%9D%B8%EA%B5%AC&aqs=chrome..69i57j46i199i465i512j0i512j46i175i199i512j0i512l5.3544j0j7&sourceid=chrome&ie=UTF-8
# pop size in North Korea = 25.97*1e6
# Population size 25970000 comes from a Google search
# 350000 already infected people on Jan
I0 <- 1e5
y0 <- c(S=25970000-I0, E=0, I=I0, R=0, CI=0)

params <- list() # input parameters for the SEIR model
# epsilon and gamma from Kim et al. (2021)
params$epsilon <- 1/3 # mean latent period = 1/epsilon
params$delta <- 1/5.2 # mean incubation period = 1/epsilon
params$gamma <- 1/6.5 # mean infectious period = 1/gamma
tend <- 50 # simulation end time 50 days
params[["R0"]] <- 3
params[["betavol"]] <- 0.4

d <- data.frame(daily_infected = dat$symptomatic)

part <- pfilter(params = params,
                  y = y0, # initial values of state variables
                  npart = 1e4,
                  data = d,
                  tend = nrow(d)) # input data set)
```


### Summarizing results
```{r}
daily_inc_summary <- t(apply(part$latent_var_filtered[,,5], 2, quantile,
            probs=c(0.025, 0.5, 0.975)))

df <- cbind(data.frame(time = dat$newdate, observed = d$daily_infected), daily_inc_summary)

ggplot(df, aes(x=time)) +
  geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), fill="steelblue", alpha=0.8)+
  geom_line(aes(y=`50%`), color="steelblue")+
  geom_point(aes(y=observed), color = "darkred")+
  labs(x="Time", y="Daily incidence")+
  theme_bw()

# ggsave("plots/daily_incidence_pf.png")

dur <- 1/params$gamma
daily_Rt_summary <- t(apply(part$beta_filtered * dur, 1, quantile,
                            probs=c(0.025, 0.5, 0.975)))  

df <- as.data.frame(daily_Rt_summary)
df$time <- dat$newdate
           
ggplot(df, aes(x=time)) +
  geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), fill="darkgreen", alpha=0.8)+
  geom_line(aes(y=`50%`), color="darkgreen")+ 
  labs(x="Time", y=expression(italic(R[t])))+
  theme_bw()

# ggsave("plots/daily_Rt_pf.png")
```

### Particle filter
```{r}
## Let's create the data

pfilter <- function (params, # parameters
                     y, # initial values of state variables
                     data, # input data set
                     npart = 1000, # number of particlar
                     tend = NULL, # simulation stop time
                     dt = 0.2,
                     observed_variable = "daily_infected",
                     modelled_variable = "CI") {
  
  # Assumptions - using daily growth rate
  nstatevar <- length(y) # number of state variables
  if(is.null(tend)) {
    tend = nrow(data)
  }
  # to store state variables a
  latent_var <- array(0,
                      dim = c(npart, tend, nstatevar),
                      dimnames = list(NULL, NULL, names(y)))
  # latent_var[, 1, ] <- y
  for (nm in names(y)) { # initial value
    latent_var[, 1, nm] <- y[[nm]]
  }
  ## parameters 
  gamma <- params[["gamma"]]
  beta0 <- params[["R0"]] * gamma
  beta_sd <- params[["betavol"]]
  beta <- matrix(rnorm(npart * tend, mean = 0, sd = beta_sd), nrow = tend)
  beta[1,] <- beta0 # this is updated at t=2
  
  wt <- matrix(NA, nrow = npart, ncol = tend) # weight (likelihood)
  wt[, 1] <- 1 / npart  # initial weights
  W <- matrix(NA, nrow = npart, ncol = tend) # normalized weights
  A <- matrix(NA, nrow = npart, ncol = tend) # Resample according to the normalized weight
  
  for (t in 2:tend) {# begin particle loop
    # beta changes according to a Geometric Brownian motion 
    beta[t, ] <- beta[t-1, ] * exp(beta[t, ])
    # run process model
    latent_var[, t, ] <- SEIR_step(params = params,
                                   y = latent_var[, t-1, ],
                                   tbegin = t-1,
                                   tend = t,
                                   dt = dt,
                                   beta = beta[t,])
    # calculate weights (likelihood)
    # wt[, t] <- assign_weights(var = latent_var, t = t, data = data)
    
    case_expected <- latent_var[, t, modelled_variable]
    case_data <- round(unlist(data[t, observed_variable]))
    expected_val <- pmax(0, case_expected) # make sure that the value is not negative
    # log_lik <- dpois(round(case_data), lambda = expected_val, log = T)
    # wt[, t] <- exp(log_lik)
    wt[, t] <- dpois(round(case_data), lambda = expected_val)
    # normalize particle weights
    W[, t] <- wt[, t] / sum(wt[, t])
    # resample particles by sampling parent particles according to weights
    A[, t] <- sample(1:npart, prob = W[1:npart, t], replace = T)
    # Resample particles for corresponding variables
    latent_var[, t,] <- latent_var[A[, t], t,]
    beta[t,] <- beta[t, A[, t]] #- needed for random walk on beta
  } # end particle loop
  
  # Marginal likelihoods
  lik_values <- rep(NA, tend)
  for (t in 1:tend) {
    lik_values[t] <- log(sum(wt[1:npart, t])) # log-likelihoods
  }# averaged log likelihoods log(L/(npart^tend))
  loglik <- - tend * log(npart) + sum(lik_values)
  
  return (list(lik_marginal = lik_values,
               lik_overall_average = loglik,
               latent_var_filtered = latent_var,
               beta_filtered = beta,
               W = W, A = A))
}
```


### SEIR step
```{r}
## differential equations are modeled using the Euler method.
## This makes it easier 
# Solve the equation using the Euler-Maruyama method
SEIR_step <- function (params = NULL,
                       y = NULL,
                       tbegin = 0,
                       tend = 1,
                       dt = 0.2,
                       beta = NULL) {
  
  y[, c("CI")] <- 0 # reset to zero to hold values from tbegin to tend
  
  S <- y[, "S"]
  E <- y[, "E"]
  I <- y[, "I"]
  R <- y[, "R"]
  daily_infected <- y[, "CI"]
  
  N <- S + E + I + R
  eps <- params[["epsilon"]]
  gam <- params[["gamma"]]
  
  for (i in seq((tbegin + dt), tend, dt)) {
    # beta is already assumed to be adjusted by N such that it can
    # be translated to Rt by multiplying the duration of infectiousness
    S_to_E <- beta * I * dt
    E_to_I <- E * eps * dt
    I_to_R <- I * gam * dt
    
    # Update state variables
    S <- S - S_to_E
    E <- E + S_to_E - E_to_I
    I <- I + E_to_I - I_to_R
    R <- R + I_to_R
    daily_infected <- daily_infected + S_to_E
  }
  
  y[, "S"] <- S
  y[, "E"] <- E
  y[, "I"] <- I
  y[, "R"] <- R
  y[, "CI"] <- daily_infected
  
  return(y)
}
```



### Run particle filter
```{r}
# North Korea data cleaned a bit
dnorth <- readRDS("outputs/NorthKorea_dat_20230122.rds")
source("src/SEPAIR_step.R")
source("src/pfilter.R")
# Initial conditions
# 350000 already infected people on Jan
I0 <- 350000
fa <- 0.306 # fraction of asymptomatic people
y0 <- c(S=25970000-I0, E=0, P=0, A=fa*I0, I=(1-fa)*I0, R=0, CE=0, CI=0)

params <- list() # input parameters for the SEIR model
# epsilon and gamma from Kim et al. (2021)
params$epsilon <- 1/3 # mean latent period = 1/epsilon
params$delta <- 1/5.2 # mean incubation period = 1/delta
params$gamma <- 1/6.5 # mean infectious period = 1/gamma
params$R0 <- 3
params$betavol <- 0.4
params$fa <- fa # fraction of asymptomatic state
params$bp <- 1 # relative infectiousness of pre-symptomatic state
params$ba <- 1 # relative infectiousness of asymptomatic state
  
# d <- data.frame(daily_symptom = dnorth$symptomatic)
d <- data.frame(daily_infected = dnorth$symptomatic)

part <- pfilter(params = params,
                y = y0, # initial values of state variables
                func = SEPAIR_step,
                npart = 1e4,
                data = d,
                tend = nrow(d), 
                observed_variable = "daily_infected", 
                modelled_variable = "CE") # input data set)
```
### Summarizing results
```{r}
library(ggplot2)
daily_inc_summary <- t(apply(part$latent_var_filtered[,,5], 2, quantile,
            probs=c(0.025, 0.5, 0.975)))

df <- cbind(data.frame(time = dnorth$newdate, observed = d$daily_infected), daily_inc_summary)

ggplot(df, aes(x=time)) +
  geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), fill="steelblue", alpha=0.8)+
  geom_line(aes(y=`50%`), color="steelblue")+
  geom_point(aes(y=observed), color = "darkred")+
  labs(x="Time", y="Daily incidence")+
  theme_bw()

# ggsave("plots/daily_incidence_pf.png")

dur <- 1/params$gamma
daily_Rt_summary <- t(apply(part$beta_filtered * dur, 1, quantile,
                            probs=c(0.025, 0.5, 0.975)))  

df <- as.data.frame(daily_Rt_summary)
df$time <- dat$newdate
           
ggplot(df, aes(x=time)) +
  geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), fill="darkgreen", alpha=0.8)+
  geom_line(aes(y=`50%`), color="darkgreen")+ 
  labs(x="Time", y=expression(italic(R[t])))+
  theme_bw()

# ggsave("plots/daily_Rt_pf.png")
```
