---
output: github_document
---


```{r global_options, echo=F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE)
```

### COVID-19 in North Korea

```{r  results='hide'}
library(readxl)
library(ggplot2)

dat <- read_xlsx("data/covid19_daily.xlsx")
names(dat) <- c("date", "symptomatic", "recovered", "deaths", "cumul_symptomatic", "cumul_recovered", "cumul_isolated", "cumul_deaths")
dat <- dat[3:(nrow(dat)-1),]
library(dplyr)
dat %>% 
  mutate(across(2:8, as.integer),
         newdate = seq(from=as.Date("2022-05-13"), 
                       to=as.Date("2022-08-10"), by="day"), 
         cfr = cumul_deaths / cumul_symptomatic) -> dat

# d <- dat[,"newdate", ]
# sum((dat$cumul_symptomatic - dat$cumul_recovered - dat$cumul_isolated) != 0)
# tstamp <- format(Sys.time(), "%Y%m%d")
# saveRDS(dat, paste0("outputs/NorthKorea_dat_", tstamp, ".rds"))
```

- 신규 유열자 (new symptomatic cases) were first repoorted on `r format(dat$newdate[1], "%B %d, %Y")` ($n$ = `r dat$symptomatic[1]`).
- There were, however, `r dat$cumul_symptomatic[1]` symptomatic cases already on `r format(dat$newdate[1], "%B %d, %Y")`.
- In total, `r dat$cumul_symptomatic[nrow(dat)]` cases were reported `r format(dat$newdate[1], "%B %d, %Y")` through `r format(dat$newdate[nrow(dat)], "%B %d, %Y")`
- Case fatality ratio was very low: CFR = `r format(round(dat$cfr[nrow(dat)], digits=8), scientific=FALSE)`

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(tidyr)

dat %>%  select(newdate, symptomatic, cumul_symptomatic) %>%
  pivot_longer(cols=c(2,3), names_to = "daily_cumul") -> d

plt <- ggplot(d, aes(newdate, value))+
  geom_line()+
  geom_point()+
  labs(x="", y="")+
  ggtitle("Symptomatic cases")+
  facet_wrap(~daily_cumul, nrow=2, scales="free_y")

plt
# Alene, Muluneh, Leltework Yismaw, Moges Agazhe Assemie, Daniel Bekele Ketema, Wodaje Gietaneh, and Tilahun Yemanu Birhan. “Serial Interval and Incubation Period of COVID-19: A Systematic Review and Meta-Analysis.” BMC Infectious Diseases 21, no. 1 (December 2021): 1–9. https://doi.org/10.1186/s12879-021-05950-x.

#Abstract: ... mean serial interval of COVID-19 was 5.2 (95%CI: 4.9–5.5) days
SI <- 5.2
SI_std <- (5.5-5.2)/1.96
```


